Formal Verification of Real-Time Data Processing of the LHC Beam Loss Monitoring System: A Case Study

### 2011-03-28 fmics/paper.tex
merge changes from dropbox

\usepackage{times,graphicx,epsfig,boxedminipage,url,xspace,array,amsmath}

@@

Our goal is to build a generic model of the SRS structure, and to prove that it satisfies its specification, that it calculates approximate running sums of received count values within acceptable error margins.
Let $\mathsf{RS}\;n$ denote, as in Figure~\ref{fig:srs}, the $n$-th output of the SRS structure, which is supposed {\bf really? supposed to have a delay?} to compute a sum of received count values with some delay, and let $\mathsf{true\_sum}\;n$ denote this sum {\bf ``this'' being the delayed one, or the non-delayed one?}.
A sketch of the desired correctness statement is \[\forall{n} \cdot \, \mathsf{RS}\;n= \mathsf{true\_sum}\; n \pm\text{acceptable error}\]
Although Figure~\ref{fig:srs} suggests that the SRS structure has only twelve outputs  (i.e. $0\leq{n}\leq11$), we obtain a more generic result (that is useful in future upgrades of the system) by proving the correctness of the above statement for all values of $n$.

@@

\begin{figure}
\caption{
Definitions of the HOL functions comprising our model of the SRS structure.
\(\mathsf{tap}\) and \(\mathsf{input}\) are defined to match Figure~\ref{fig:srs}.
Slice $0$ is a fake slice representing the SRS input; this enables a succinct definition of $\mathsf{source}$.
\label{fig:definitions}}
\begin{align*}
\mathsf{tap}\;0\;0&=0&\mathsf{tap}\;0\;x&=0\\
\mathsf{tap}\;1\;0&=1-1&\mathsf{tap}\;1\;x&=2-1\\
\mathsf{tap}\;2\;0&=8-1&\mathsf{tap}\;2\;x&=16-1\\
&\dots&\mathsf{tap}\;6\;x&=128-1\quad\text{(where each $x>0$)}\\\\
\mathsf{input}\;0&=(0,0)&\mathsf{input}\;1&=(0,0)\\
\mathsf{input}\;2&=(0,0)&\mathsf{input}\;3&=(1,1)\\
\mathsf{input}\;4&=(3,0)&\mathsf{input}\;5&=(4,0)\\
\mathsf{input}\;6&=(4,1)&\mathsf{input}\;n&=(n-1,0)\quad\text{(where $n>6$)}
\end{align*}
\begin{align*}
\mathsf{delay}\;0&=1\\
\mathsf{delay}\;n&=\mathsf{delay}\;n'\times((\mathsf{tap}\;n'\;x)+1)\\
&\qquad\text{where $(n',x)=\mathsf{input\;n}$ (and $n>0$)}
\end{align*}
\begin{align*}
\mathsf{update\_time}\;n\;t&\iff(t\operatorname{mod}\mathsf{delay}\;n=0)
\end{align*}
\begin{align*}
\mathsf{source}\;D\;n\;0\;t&=\mathsf{output}\;D\;n'\;x\;t\quad\text{where $(n',x)=\mathsf{input\;n}$}\\
\mathsf{source}\;D\;n\;m\;t&=\;\mathsf{SR}\;D\;n\;(m-1)\;\quad\text{(where $m>0$)}
\end{align*}
\begin{align*}
\mathsf{output}\;D\;0\;x\;t&=D\;t\\
\mathsf{output}\;D\;n\;x\;0&=0\\
\mathsf{output}\;D\;n\;x\;t&=
\text{if $\mathsf{update\_time}\;n\;t$}\\
&\qquad\text{then }((\mathsf{output}\;D\;n\;x\;(t-1))+(\mathsf{source}\;D\;n\;0\;(t-1)))-(\mathsf{SR}\;D\;n(\mathsf{tap}\;n\;x)\;(t-1))\\
&\qquad\text{else }\mathsf{output}\;D\;n\;x\;(t-1)\qquad\text{(where $t>0$)}
\end{align*}
\begin{align*}
\mathsf{SR}\;D\;n\;m\;0&=0\\
\mathsf{SR}\;D\;n\;m\;t&=\text{if $\mathsf{update\_time}\;n\;t$}\\
&\qquad\text{then }\mathsf{source}\;D\;n\;m\;(t-1)\\
&\qquad\text{else }\mathsf{SR}\;D\;n\;m\;t\qquad\text{(where $t>0$)}
\end{align*}
\begin{align*}
\mathsf{RS}\;D\;n\;t&=\mathsf{output}\;D\;\left(\left\lfloor\frac{n}{2}\right\rfloor+1\right)\;(n\operatorname{mod}2)
\end{align*}
\end{figure}

@@

>>>>

\usepackage{times,graphicx,epsfig,boxedminipage,url,xspace,array,amsmath}
\newcommand\T{\rule{0pt}{2.6ex}}
\newcommand\B{\rule[-1.2ex]{0pt}{0pt}}

@@

Our goal is to build a generic model of the SRS structure, and to prove that it satisfies its specification, that it calculates approximate running sums of received count values within acceptable error margins.
Let $\mathsf{RS}\;n$ denote, as in Figure~\ref{fig:srs}, the $n$-th output of the SRS structure, which is supposed to compute a sum of received count values, and let $\mathsf{true\_sum}\;n$ denote this sum.
The multi-layerd structure of the SRS and the read delay of each shift register result in the outputs of the SRS to be delayed from the $\mathsf{true\_sum}$ values.
A sketch of the desired correctness statement is \[\forall{n} \cdot \, \mathsf{RS}\;n= \mathsf{true\_sum}\; n \pm\text{acceptable error}\]
Although Figure~\ref{fig:srs} suggests that the SRS structure has only twelve outputs  (i.e. $0\leq{n}\leq11$), we obtain a more generic result (that is useful in future upgrades of the system) by proving the correctness of the above statement for all values of $n$.

@@

\begin{figure}
\caption{
Definitions of the HOL functions comprising our model of the SRS structure.
\(\mathsf{tap}\) and \(\mathsf{input}\) are defined to match Figure~\ref{fig:srs}.
Slice $0$ is a fake slice representing the SRS input; this enables a succinct definition of $\mathsf{source}$.
\vspace{0.2in}
\label{fig:definitions}}
\begin{tabular}{l|ll}
Function \T \B & Definition \\\hline
$\mathsf{tap}\;n\;x$ \T & $\;\mathsf{tap}\;0\;0=0\quad\quad$ & $\mathsf{tap}\;0\;x=0$\\
&$\;\mathsf{tap}\;1\;0=1-1\quad$ &  $\mathsf{tap}\;1\;x=2-1$\\
&$\;\mathsf{tap}\;2\;0=8-1\quad$ &  $\mathsf{tap}\;2\;x=16-1$\\
&$\;\dots\quad$ \B & $\mathsf{tap}\;6\;x=128-1\quad\text{(where each $x>0$)}$\\\hline

$\mathsf{input}\;n$ \T & $\;\mathsf{input}\;0=(0,0)\quad$ & $\mathsf{input}\;1=(0,0)$\\
&$\;\mathsf{input}\;2=(0,0)\quad$ & $\mathsf{input}\;3=(1,1)$\\
&$\;\mathsf{input}\;4=(3,0)\quad$ & $\mathsf{input}\;5=(4,0)$\\
&$\;\mathsf{input}\;6=(4,1)\quad$ \B & $\mathsf{input}\;n=(n-1,0)\quad\text{(where $n>6$)}$\\\hline

$\mathsf{delay}\;n$ \T & $\;\mathsf{delay}\;0=1$\\
&\multicolumn{2}{|l}{$\;\mathsf{delay}\;n=\mathsf{delay}\;n'\times((\mathsf{tap}\;n'\;x)+1)$}\\
&\multicolumn{2}{|l}{$\qquad\qquad\qquad\text{where $(n',x)=\mathsf{input\;n}$ (and $n>0$)}$} \B \\\hline

$\mathsf{source}\;D\;n\;m\;t$ \T & \multicolumn{2}{|l}{\;$\mathsf{source}\;D\;n\;0\;t=\mathsf{output}\;D\;n'\;x\;t\quad\text{where $(n',x)=\mathsf{input\;n}$}$}\\
&\multicolumn{2}{|l}{$\;\mathsf{source}\;D\;n\;m\;t=\;\mathsf{SR}\;D\;n\;(m-1)\;\quad\text{(where $m>0$)}$} \B \\\hline

$\mathsf{SR}\;D\;n\;m\;t$ \T &  \multicolumn{2}{|l}{\;$\mathsf{SR}\;D\;n\;m\;0=0$}\\
&\multicolumn{2}{|l}{$\;\mathsf{SR}\;D\;n\;m\;t=\text{if $\mathsf{update\_time}\;n\;t$}\;\;\;\text{then }$}\\
&&$\mathsf{source}\;D\;n\;m\;(t-1)$\\
&\multicolumn{2}{|l}{$\qquad\qquad\qquad\quad\text{else }\mathsf{SR}\;D\;n\;m\;t\qquad\text{(where $t>0$)}$}\B \\\hline

$\mathsf{output}\;D\;n\;x\;t$ \T &\multicolumn{2}{|l}{$\;\mathsf{output}\;D\;0\;x\;t=D\;t$}\\
&\multicolumn{2}{|l}{$\;\mathsf{output}\;D\;n\;x\;0=0$}\\
&\multicolumn{2}{|l}{$\;\mathsf{output}\;D\;n\;x\;t = \text{if $\mathsf{update\_time}\;n\;t$}\;\;\;\text{then }$}\\
&\multicolumn{2}{|l}{$\qquad\quad((\mathsf{output}\;D\;n\;x\;(t-1))+(\mathsf{source}\;D\;n\;0\;(t-1)))-$}\\
&&$\qquad\qquad\qquad(\mathsf{SR}\;D\;n(\mathsf{tap}\;n\;x)\;(t-1))$\\
&\multicolumn{2}{|l}{$\qquad\qquad\qquad\qquad\text{else }\mathsf{output}\;D\;n\;x\;(t-1)\qquad\text{(where $t>0$)}$}\B\\\hline

$\mathsf{RS}\;D\;n\;t$ \T \B & \multicolumn{2}{|l}{$\;\mathsf{RS}\;D\;n\;t=\mathsf{output}\;D\;\left(\left\lfloor\frac{n}{2}\right\rfloor+1\right)\;(n\operatorname{mod}2)$}\\\hline

$\mathsf{update\_time}\;n\;t$ \T & \multicolumn{2}{|l}{$\;\mathsf{update\_time}\;n\;t\iff(t\operatorname{mod}\mathsf{delay}\;n=0)$}
\end{tabular}
\end{figure}

@@

%One of our objectives in formally modelling SRS was to keep the topology parameterized.
%It is likely that the configuration of the slices and shift registers will in the future change to accommodate higher precision.
%A parameterized model can be used both for current and future versions of the system.
%For this, we formalized the definitions of the model by six parameters:

%Our work proceeded in two phases, starting with a model of a simplified structure and proceeding to a model which is closer to the SRS implementation in BLMS.
%In both phases, we model each building block of the SRS structure that holds a count value -- for example, each cell in each shift register -- as a function in HOL that takes the input stream and current time step as inputs and returns the value the corresponding building block holds at that time.

%Our simplifications in the first phase include:
%\begin{enumerate}
%\item Masking the middle taps in each slice,
%\item Considering a more regular arrangement of slices than the SRS implementation in BLMS,
%\item Masking the subtractors and the accumulators in each slice, instead, defining the output of a slice as the sum of the values in its cells directly,
%\item Simplifying timing by ignoring some of the delays that each building block may introduce, {\bf this is not true - there were delays in the simplified model}
%\item Using natural numbers, rather than finite words, for count values.
%\end{enumerate}

%Building a logical model for simplified structure of the SRS helped us improve our understanding of the behavior of the SRS.

%In the second phase, we eliminated the first four simplifications (listed above) by adding middle taps, a more flexible arrangement of slices, the subtractor and accumulator, and considering the delays that each block may introduce.
%In both cases, however, the model is simultaneously a simplification and a generalization of the actual structure of SRS in BLMS in the following way: we model an infinite number of slices (rather than six slices), each with an infinite number of shift register cells and taps (rather than fixed width shift registers and only two taps), simply by letting indices range over the natural numbers without explicitly giving limits.
%This parameterization makes the model and the analysis more likely to be applicable to future upgraded/modified versions of the system.
